
\documentclass{sig-alternate}
\usepackage{balance,array,microtype,pifont}
\usepackage[bf]{subfigure}

\let\proof\relax
\let\endproof\relax


\usepackage{color, times, mathptmx, amsmath, amssymb, amsthm, verbatim, 
  mdwlist, graphicx, footnote, multirow, xspace, tikz, colortbl,dcolumn, mathtools,zi4,listings}


\usepackage[noadjust]{cite}
\usepackage{changepage}

\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\let\footnotesize\small

\usepackage[boxed]{algorithm}% http://ctan.org/pkg/algorithms
\usepackage[noend]{algpseudocode}% http://ctan.org/pkg/algorithmicx


\setlength{\parskip}{3ex plus 2ex minus 2ex}

\theoremstyle{definition}
\setlength{\textfloatsep}{1em}% Remove \textfloatsep

\renewcommand{\floatpagefraction}{.99}
\renewcommand{\topfraction}{.99}
\renewcommand{\bottomfraction}{.99}

\interfootnotelinepenalty=10000

\usepackage{scrextend}
\usepackage[hyphens]{url}
\usepackage{breakurl}

\newcommand{\commentt}[1]{{\small\texttt{#1}}}

\newcommand{\miniheadnostop}[1]{{\vspace{.4em}\noindent\textit{#1} }}
\newcommand{\minihead}[1]{{\vspace{.45em}\noindent\textbf{#1.} }}
\newcommand{\miniheadit}[1]{{\vspace{.45em}\noindent\textit{#1.} }}
\newcommand{\miniheadnopd}[2]{{\vspace{.4em}\noindent\textbf{#1}\hspace{.5em}{#2}\vspace{.4em} }}
\newcommand{\minidef}[1]{{\vspace{.25em}\noindent\textit{#1} }}

\newcommand{\pbnote}[1]{{\color{red}{#1 ---P.B.}}}

\renewcommand{\ttdefault}{zi4}
\usepackage[T1]{fontenc}

\algblock{ParFor}{EndParFor}
% customising the new block
\algnewcommand\algorithmicparfor{\textbf{parfor}}
\algnewcommand\algorithmicpardo{\textbf{do}}
\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
\algrenewtext{EndParFor}{\vspace{-1em}}


\newenvironment{myitemize}
{
  \vspace{-.4em}
    \begin{list}{$\bullet$ }{}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.4em}
}

\newenvironment{introenumerate}
{

   \vspace{-.5em}
   \newcounter{qdcounter}
    \begin{list}{\arabic{qdcounter}.~}{\usecounter{qdcounter}\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.5em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}


\newenvironment{myenumerate}
{

   \vspace{-.5em}
   \newcounter{qdcounter}
    \begin{list}{\arabic{qdcounter}.~}{\usecounter{qdcounter}\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.5em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}


\usepackage[leftmargin=0em]{quoting}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{XXX}{YYY}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Feral Concurrency Control: A Study of\\Application Integrity in
  a Modern Web Framework}

{\author{}}
\maketitle

\section*{TODO}
\begin{myitemize}
\item Section~\ref{sec:rails-cc}: Add statistics on when optimistic
  locking is used.
\item Section~\ref{sec:rails-cc}: Add discussion about how often locks
  are used.
\item Section~\ref{sec:apps}: Add pseudocode for how validation is performed.
\item Section~\ref{sec:rails-cc}: Expand discussion of which apps use
  which mechanisms. Specifically, who uses transactions and why?
\item Section~\ref{sec:evaluation}: Add pseudocode for sample applications.
\item Section~\ref{sec:evaluation}: Add pseudocode for evaluation
  workload.
\item Section~\ref{sec:evaluation}: Add discussion of how association
  checking can go wrong.
\item Section~\ref{sec:evaluation}: Add discussion of worst case
  degree of inconsistency for association races.
\item Section~\ref{sec:other-orms}: Expand discussion of what these
  frameworks actually do; too terse right now as to be ambiguous.
\item Section~\ref{sec:other-orms}: Figure out if waterline has custom
  validations.
\item Section~\ref{sec:other-orms}: Figure out if Django backs custom
  validations with a transaction.
\item Section~\ref{sec:conclusion}: Add conclusion
\item General: convert some of these footnotes to references
\end{myitemize}

\input{abstract}

\input{intro}

\input{motivation}

\input{feral}

\input{applications}

\input{evaluation}

\input{discussion}

\input{relatedwork}

\input{conclusion}

%\scriptsize
%\linespread{.98}
%\selectfont


\bibliography{feral-cc} \bibliographystyle{abbrv}

\input{appendix}

\balance


\end{document}
