
\section{Related Work}
\label{sec:relatedwork}

There is a large body of related work that we consider in four
categories: object relational mapping systems, the study of weak
isolation and applications requirements, the quantification of isolation
behavior, and empirical open source software analysis.

\minihead{ORMs} Database systems and application programming
frameworks have a long
history~\cite{objectstore,shore,bernstein-orm}. The ``impedance
mismatch'' between object-oriented programming and the relational
model is a perennial problem in data management systems. Ruby on Rails
is no exception, and the concurrency control issues we study here are
endemic to this mismatch---namely, disuse of common concurrency
control mechanisms like database-backed constraints. Bridging this gap
remains an active area of research~\cite{db-to-model}.

The latest wave of web programming frameworks has inspired diverse
research spanning databases, verification, and security. StatusQuo
uses program analysis and synthesis to transform imperative ORM code
into SQL, leveraging the efficiency of database-backed web
applications written in the Spring framework~\cite{statusquo}. Rails
has been the subject of study in the verification of cross-site
scripting attacks~\cite{rails-xss}, errors in data
modeling of associations~\cite{rails-bounded}, and arbitrary,
user-specified (non-validation) invariants~\cite{invariant-web}.
Rails-style ORM validations have been used to improve systems security
via client-side execution~\cite{waves,caveat}. Our focus here is on
the concurrency control requirements and usages of applications
written in Rails.

\minihead{Applications and weak isolation} The issues we examine here
are fundamental to the use of weak isolation in data management
systems. Non-serializable isolation dates to the
mid-1970s~\cite{gray-isolation} and has a colorful
history~\cite{adya-isolation}; today, by volume, many database
management systems are non-serializable by
default~\cite{hat-vldb}. The isolation anomalies surfaced by
the stores we study here are directly responsible for violating the
integrity of the validations we consider.

However, serializable isolation is not strictly necessary for
maintaining application integrity. Semantic-based concurrency
control criteria has almost as long a lineage as
serializability~\cite{eswaran-consistency,ic-survey-two} and suggests
that, with additional, non-syntactic knowledge about applications
(e.g., integrity constraints)~\cite{kung1979optimality}, correctness
is achievable without serializability. This use of invariants has
enjoyed recent popularity in work by Li et al.~\cite{redblue-new}, Roy
et al.~\cite{writes-forest}, and Bailis et al.~\cite{coord-avoid}. We
use the concept of invariant confluence from~\cite{coord-avoid} to
determine whether Rails's built-in validators and applications written
in Rails are indeed safe under any coordination-free execution. Our
methodology is closest in spirit to~\cite{coord-avoid}, but, here, we
examine real applications instead of standardized benchmarks.

\minihead{Quantifying anomalies} A range of research similarly
quantifies the effect of non-serializable isolation in a variety of
ways.

Perhaps closest to our work is a study by Fekete et al., which
quantitatively analyzed data inconsistencies arising from
non-serializable schedules~\cite{fekete-quantifying}. This study used
a hand-crafted benchmark for analysis but is nevertheless one of the only
studies of actual application inconsistencies. Here, we focus on open
source applications from the Rails community.

A larger body of work examines isolation anomalies at the read-write
interface (that is, measures deviations from properties such as
serializability or linearizability but \textit{not} the end effect of
these deviations on actual application behavior). Wada et
al. evaluated the staleness of Amazon's SimpleDB using end-user
request tracing~\cite{wada-data}, while Bermbach and Tai evaluated
Amazon S3~\cite{bermbach-eventual}, each quantifying various forms of
non-serializable behavior. Golab et al. provide algorithms for
verifying the linearizability of and sequential consistency arbitrary
data stores~\cite{golab-analyzing} and Zellag and Kemme provide
algorithms for verifying their
serializability~\cite{zellag-consistent} and other cycle-based
isolation anomalies~\cite{zellag-real}. Probabilistically Bounded
Staleness provides time- and version-based staleness predictions for
eventually consistent data stores~\cite{pbs}. Our focus here is on
anomalies as observed by application logic rather than read-write
anomalies observed under weak isolation.

\minihead{Empirical software analysis} Empirical software analysis of
open source software is a topic of active interest in the software
engineering research community~\cite{foss-icse}. In the parlance of
that community, in this work, we perform a mixed-methods analysis,
combining quantitative survey techniques with a confirmatory case
study of Rails's susceptibility to validation
errors~\cite{empiricalmethods}. In our survey, we attempt to minimize
sampling bias towards validation-heavy projects by focusing our
attention on popular projects, as measured by GitHub stars. Our use of
quantitative data followed by supporting qualitative data from
documentation and issue tracking---as well as the chronology of
methodologies we employed to attain the results presented here---can
be considered an instance of the sequential exploration
strategy~\cite{creswell2013research}. We specifically use these techniques in
service of better understanding use of database
concurrency control.
